---
title: "Anomaly Detection"
author: "Jenipher Mawia"
date: "11/13/2020"
output: html_document
---
identify unexpected data points in unlabeled datasets

# 1. Problem Definition

- Detect anomalies in the sales data.

- Implement the solution using unsupervised learning techniques for anomaly detection

# 2. Defining the metrics for success
This project will be considered a success if anomalies that indicate fraud are detected. 


# 3. The Context
You are a Data analyst at Carrefour Kenya and are currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax). The main task here is to detect fraud in any of the sales made. 

# 4. Experimental Design taken
The following is the order in which I went about to achieve the objectives of this project: 

- Data Sourcing and Understanding
- Checking the data (head, shape(number of records), datatypes)
- Data Type Conversion
- Anomaly detection
- Plotting anomalies

# 5. Data Sourcing and Understanding. 
The data was provided by Moringa School and can be downloaded [here](http://bit.ly/CarreFourSalesDataset)


```{r}
# load libraries
library(tidyverse)
library(tibbletime)
library(anomalize)
```

## Reading the data 

```{r}
sales_data <- read.csv("http://bit.ly/CarreFourSalesDataset")
```

# 6. Checking the data

checking the top of the data

```{r}
head(sales_data)
```

Checking the data types

```{r}
str(sales_data)
```

Checking the dimensions of the data

```{r}
dim(sales_data)
```
There are two columns and 1000 entries in the data of types character and numeric. 

The date column shoudl be converted to its appropriate format. 

# 7. Data Type Conversion

### Change the data type of the "Date" column 
```{r}
# convert the datatype of the date column
sales_data$Date <- anytime::anydate(sales_data$Date)
str(sales_data)
```

Extract the year, month and day from the date column

```{r}
sales_data$year <- as.numeric(format(sales_data$Date, format="%Y"))
sales_data$month <- as.numeric(format(sales_data$Date, format="%m"))
sales_data$day <- as.numeric(format(sales_data$Date, format="%d"))
#check the datatypes
head(sales_data)

```


```{r}
# change the class of the data to tibble to be able to use the anomalize function
anomaly <- as_tibble(sales_data)
head(anomaly)
```

# 8. Checking for anomalies

```{r}
# check for anomalies in the Sales column using the anomalize() function
sales_anomalies <- anomalize(anomaly, Sales, method = "iqr", alpha = 0.1, max_anoms = 0.2, verbose = FALSE)

head(sales_anomalies)
```

```{r}
# check for unique values in the column "anomaly"
unique(sales_anomalies$anomaly)
```

Some transactions have been detected as anomalies in the data. These could be signs of fraud.  

### Converting the anomaly column to factor

```{r}
# convert the anomaly column to factor
sales_anomalies_data<- sales_anomalies
sales_anomalies_data$anomaly <- as.factor(sales_anomalies_data$anomaly)
str(sales_anomalies_data)
```

# 9. Plotting the anomalies

```{r}
#plot(sales_anomalies_data$Sales, color = "anomaly")

sales_anomalies_data %>% 
  ggplot() +
  aes(x = Sales, anomaly = ..count../nrow(sales_anomalies_data), fill = anomaly) +
  geom_dotplot() +
  ylab("relative frequency")
```

From the plot we can see that anomalies were detected at sales greater than 1000
         
