---
title: "Association Rules"
author: "Jenipher Mawia"
date: "11/13/2020"
output:
  word_document: default
  html_document: default
---

```{r}
# load libraries
library(arules)
```

# Reading the data

```{r}
# since we want the data as class transactions, we will read the data using read.transactions function
association <- read.transactions("http://bit.ly/SupermarketDatasetII")

```

# Checking the data

```{r}
association
```

There are 7501 transactions and 5729 items in the data

# Verifying the object's class

```{r}
# checking the class of our transactions data
class(association)

```

This shows us transactions as the type of data that we will need

# Data Exploration

```{r}
# Previewing the first 5 transactions
inspect(association[1:5])
```


```{r}
# preview the items that make up our dataset,
items<-as.data.frame(itemLabels(association))
colnames(items) <- "Item"
head(items,15)

```


```{r}
# Generating a summary of the transaction dataset
# ---
# This would give us some information such as the most purchased items, 
# distribution of the item sets (no. of items purchased in each transaction), etc.
# ---
# 
summary(association)
```


```{r}
# Exploring the frequency of some articles 
# i.e. transacations ranging from 12 to 16 and performing 
# some operation in percentage terms of the total transactions 
# 
itemFrequency(association[, 12:16],type = "absolute")
round(itemFrequency(association[, 12:16],type = "relative")*100,2)
```


```{r}
# Producing a chart of frequencies and fitering 
# to consider only items with a minimum percentage 
# of support/ considering a top x of items
# ---
# Displaying top 10 most common items in the transactions dataset 
# and the items whose relative importance is at least 10%
# 
par(mfrow = c(1, 2))

# plot the frequency of items
itemFrequencyPlot(association, topN = 10,col="darkgreen")
itemFrequencyPlot(association, support = 0.03,col="darkred")
```

```{r}
# Create an item frequency plot for the top 20 items
if (!require("RColorBrewer")) {
  # install color package of R
install.packages("RColorBrewer")
#include library RColorBrewer
library(RColorBrewer)
}
itemFrequencyPlot(association,topN=20,type="absolute",col=brewer.pal(8,'Pastel2'), main="Absolute Item Frequency Plot")
```

```{r}
itemFrequencyPlot(association,topN=20,type="relative",col=brewer.pal(8,'Pastel2'),main="Relative Item Frequency Plot")

```



```{r}
# Building a model based on association rules 
# using the apriori function 
# ---
# We use Min Support as 0.001 and confidence as 0.8
# ---
# 
rules <- apriori (association, parameter = list(supp = 0.001, conf = 0.8))
rules

```

Since there are 49122 rules, let's print only top 10:

```{r}
summary(rules)
```


```{r}
inspect(rules[1:10])
```

# Limiting the number and size of rules 

```{r}
# We use measures of significance and interest on the rules, 
# determining which ones are interesting and which to discard.
# ---
# However since we built the model using 0.001 Min support 
# and confidence as 0.8 we obtained 410 rules.
# However, in order to illustrate the sensitivity of the model to these two parameters, 
# we will see what happens if we increase the support or lower the confidence level
# 

# Building a apriori model with Min Support as 0.002 and confidence as 0.8.
rules2 <- apriori (association,parameter = list(supp = 0.002, conf = 0.8)) 

# Building apriori model with Min Support as 0.002 and confidence as 0.6.
rules3 <- apriori (association, parameter = list(supp = 0.001, conf = 0.6)) 

rules2

rules3

```

# Removing redundant rules


```{r}
subset.rules <- which(colSums(is.subset(rules, rules)) > 1) # get subset rules in vector
length(subset.rules)  #> 3913
```

```{r}

subset.association.rules. <- rules[-subset.rules] # remove subset rules
subset.association.rules.
```

# Finding Rules related to given items

```{r}
# If we're interested in making a promotion relating to the sale of yogurt, 
# we could create a subset of rules concerning these products 
# ---
# This would tell us the items that the customers bought before purchasing yogurt
# ---
# 
yogurt_rules <- apriori(association, parameter = list(supp=0.01, conf=0.8), appearance = list(default="lhs", rhs="yogurt"))

inspect(head(yogurt_rules))

# Then order by confidence
#yogurt<-sort(yogurt, by="confidence", decreasing=TRUE)
#inspect(yogurt[1:5])
```

```{r}

yogurt_rules_lhs <- apriori(association, parameter = list(supp=0.01, conf=0.8), appearance = list(default="rhs", lhs="yogurt"))
inspect(head(yogurt_rules_lhs))

```

```{r}

# build rules for the most frequent/popular items
```


# Visualizing Association Rules

```{r}
library(arulesViz)
# Filter rules with confidence greater than 0.4 or 40%
subRules<-rules[quality(rules)$confidence>0.4]
#Plot SubRules
plot(subRules)
```


```{r}
plot(subRules,method="two-key plot")

```



```{r}
# plotly_arules(subRules)



```


```{r}
#top10subRules <- head(subRules, n = 10, by = "confidence")

#plot(top10subRules, method = "graph",  engine = "htmlwidget")

```

# Individual Rule Representation


```{r}
# Filter top 20 rules with highest lift
subRules2<-head(subRules, n=20, by="lift")
plot(subRules2, method="paracoord")

```

 Finish up in the morning 4am tommorrow





































































